Proposal

### 2022-06-08 proposal/proposal.qmd
fixed bullets

---
title: "Proposal"
subtitle: "STA 210 - Project"
author: "Bayes' Harem - Christina Wang, Kat Cottrell, David Goh, Ethan Song"
format: pdf
editor: source
---

@@

Our project is on the topic of abortion attitudes in the US. Abortion has long been a divisive issue in the US, with both pro-choice and pro-life groups organizing in states across the country in support of political change that supported their respective sides (Ziegler, 2020). However, following the Supreme Court’s decision in Roe v. Wade in 1973, the issue has risen in political salience. Both the pro-choice and pro-life movement have gained national prominence and the two major political parties have polarized around the issue, with the Democratic Party in favor of and the Republican Party against policies legalizing and increasing access to abortion (Weinberger 2022). Abortion has also increasingly become a key issue that voters consider when they decide on candidates to vote for, with an increasing share of Americans becoming “single-issue voters” regarding abortion (Brenan 2020).

@@

In this project, we are researching the factors that influence Americans’ abortion attitudes. Specifically, we would like to know which demographic factors (such as age, gender, and education level) and personal attitudes towards other issues (such as political ideology, importance of religion, and respect for authority) best predict an individual’s attitude towards abortion. Given that this issue has become highly polarized by political party, we predict that holding liberal political ideology and liberal attitudes on other issues, as well as being younger in age is correlated with believing that abortion is more justified. (the exact predictors we will use will be decided after model selection).

@@

Ziegler, M. (2020, October 22). Abortion politics polarized before Roe. When it’s gone, the fighting won’t stop. The Washington Post. https://www.washingtonpost.com/outlook/2020/10/22/roe-polarize-abortion-politics/

@@

\[Kat Cottrell\]

>>>>

---
title: "Proposal"
subtitle: "STA 210 - Project"
author: "Bayes' Harem - Christina Wang, Kat Cottrell, David Goh, Ethan Song"
format: pdf
editor: source
editor_options: 
  markdown: 
    wrap: 72
---

@@

Our project is on the topic of abortion attitudes in the US. Abortion
has long been a divisive issue in the US, with both pro-choice and
pro-life groups organizing in states across the country in support of
political change that supported their respective sides (Ziegler, 2020).
However, following the Supreme Court's decision in Roe v. Wade in 1973,
the issue has risen in political salience. Both the pro-choice and
pro-life movement have gained national prominence and the two major
political parties have polarized around the issue, with the Democratic
Party in favor of and the Republican Party against policies legalizing
and increasing access to abortion (Weinberger 2022). Abortion has also
increasingly become a key issue that voters consider when they decide on
candidates to vote for, with an increasing share of Americans becoming
"single-issue voters" regarding abortion (Brenan 2020).

@@

In this project, we are researching the factors that influence
Americans' abortion attitudes. Specifically, we would like to know which
demographic factors (such as age, gender, and education level) and
personal attitudes towards other issues (such as political ideology,
importance of religion, and respect for authority) best predict an
individual's attitude towards abortion. Given that this issue has become
highly polarized by political party, we predict that holding liberal
political ideology and liberal attitudes on other issues, as well as
being younger in age is correlated with believing that abortion is more
justified. (the exact predictors we will use will be decided after model
selection).

@@

Ziegler, M. (2020, October 22). Abortion politics polarized before Roe.
When it's gone, the fighting won't stop. The Washington Post.
https://www.washingtonpost.com/outlook/2020/10/22/roe-polarize-abortion-politics/

@@

The dataset observes "attitudes on the justifiability of abortion in the
United States across six waves of World Values Survey data" (README.md)
and some basic qualities of the respondants.

Observations include:

-   WVS country code

-   Generational wave (1982, 1990, 1995, 1999, 2006, or 2011)

-   Justifiability of abortion (1-10)

-   Age (17 to 96)

-   College graduate (1 for yes)

-   Female (1 for women) - Unemployed (1 = currently unemployed)

-   Ideology (1-10 for left-right)

-   Financial satisfaction (1-10 for least-most)

```{=html}
<!-- -->
```
-   WVS post-materialist index (-1 = materialist. 2 = mixed. 3 =
    post-materialist)

-   Child autonomy index (-2 to 2 for obedience and religious
    faith-determination and independence)

-   Trust (1 = believes most people can be trusted)

-   Importance of God (1-10)

-   Opinion of respect for authority (-1-1 for bad to good)

-   National pride (1 = very proud to be an American)

The data were collected as part of the World Values Survey, which is
administered every few years and collects information about people's
values and beliefs worldwide. The survey aims to get a nationally
representative sample of a minimum of 1200 for most countries, and the
data are collected via face-to-face interviews at the respondents'
homes. The data included in this set specifically include responses from
6 waves of the survey (administered over the period 1982-2011). The
responses included in this set are from people in the United States, and
it examines their attitudes towards abortion.
